\begin{acknowledgments}
  I would first like to thank my thesis supervisors, Assistant Professor
  Yair Zick and Dr Alan Tsang of the School of Computing at the National
  University of Singapore (NUS).
  Professor Zick brought forward the original idea behind this thesis but
  consistenly encouraged this thesis to be my own work.
  He offered his invaluable domain knowledge on game theory throughout the
  research process.
  His detailed review and comments on my writing allowed me to put together
  a thesis I am truely proud of.
  I'm grateful that Dr Tsang always found the time and devoted the effort in
  exploring the ``deep ends'' and ``rabbit holes'' with me during the
  experiment phase of this research project.
  His presence and patience made the darkest moments in the journey much more
  tolerable.
  Not only he offered me timely feedback on my writing, he also provided me
  with test cases and conducted code review on the key parts of the code base.

  I would like to express my appreciation to Dr Omer Lev of the Department of
  Industrial Engineering and Management at the Ben-Gurion University of the
  Negev.
  His expertise in Israeli politics is instrumental in the qualitative analysis
  part of this project.
  I shall also acknowledge that he authored the first draft of the political
  analysis section in a paper version of this project, of which I have adopted
  and expanded in this thesis.

  I would also like to thank Tohar Lukov and Ewelina Marcinkiewicz of NUS for
  their coding work that downloads the necessary data from the Knesset website.

  I would like to extend my gratitude to the Knesset data team members who
  thoroughly responded to our queries, which allowed us to better understand
  various aspects of the data and their potential impact on our experiment.

  Last but not least, I would like to thank my family for their unconditional
  love and support throughout the process.
  I want to thank my baby Ada for boosting my productivity and efficiency, as
  her existance greatly increased the opportunity cost of wasting time.
  I want to thank my husband Tim for his continuous emotional support, as well
  as taking over tasks and responsibilities in family life which freed me more
  time to focus on my research.
  I want to thank my mother for flying, under very short notice, from China to
  Singapore to take care of Ada and myself when we both fell sick.
  This work would not have been possible without you. Thank you!

\end{acknowledgments}
