\chapter{Conclusion}
\label{ch:concl}

In this thesis we set out to group Israeli parliament members
based on incomplete preferences derived from their votes on legislative bills.
We take a data-driven approach while accounting for strategic behavior of
parliament members as players of hedonic coalition formation games.
By comparison, the clustering and community detection methods commonly used in
the machine learning/AI community do not consider players' preferences and
strategic behavior.
On the other hand, the game theory and computational social choice community has primarily
studied the theoretical aspect of hedonic coalition formation games;
the limited number of studies on the application of hedonic games are
``prescriptive'' and often use only simulated data.
Our study bridges the gap through conducting a ``descriptive'' study of hedonic
game theoretical models using real-world data of scale and with ground truth ---
the Knesset voting data on over 7000 legislative bills with party affiliations as
the baseline partition.

We demonstrate that the top three performing PAC models, according to the
information theoretical measures against party affiliations, are not only able
to recover the overall structure of government vs. opposition groups, but also
are more coherent compared to machine learning models such as $k$-means and
Stochastic Block Models.
By comparing the PAC and full-information settings for the friends models, we
observe that through sampling, PAC models produce partition results that are
more robust --- sampling helps ``dampen'' the model's sensitivity to the
definition of friends.

One limitation of our study is that it is a case study based off a single
dataset; it is only the first step, taking the empirical approach, to
demonstrate the efficacy of PAC hedonic game models.
A natural next step is to apply the top performing models, namely selective
friends and selective enemies models on other parliaments, and evaluate if the
model partitions distinguish the left wing from the right wing, the government
from the opposition parties.
A good candidate is the Dutch parliament which also has a coalition government. 
The Brazilian parliament data is public accessible\footnote{The data API is at
\url{https://dadosabertos.camara.leg.br/swagger/api.html}}, and is studied by
\citenameyear{Peixoto} using the stochastic block model, which can be compared
directly with the partitions produced by the PAC models.
The United States congress could also be interesting as one could test if the
model reveals any regional effect.
Another research direction is to apply other hedonic uncertainty models, such
as the reinforcement learning model by \citenameyear{Chalkiadakis2004} for
finding a Bayesian core, on the Knesset dataset, and compare the resulting
partitions with that of our PAC models using the same criteria.
